Network Trait marker for working memory deficit in Schizophrenia: a multi site study

### 2016-02-18 manuscript.tex
0218
0218

Working memory is a temporal process to maintain and manipulate information over short periods of time. WM has been shown to be impaired in many neurological and psychiatric syndromes including Schizophrenia. It was thought to be tightly associated with cognitive deficits in Schizophrenia (Green 2000). WM itself is a complex process which has been multidimensionally related to the psychosis.

Thus, a examination of the underlying neurological mechanism provided us with insight into the causes, progression and even treatment of schizophrenia. There has been abundant imaging studies in this area, but the differences among the tasks and material make it difficult to make a conclusion. There have been evidence of multi-modality deficits: visual spatial, visual object, verbal, and other types of working memory. One popular and original model of WM is a system encoding and maintaining modality-specific short-term memory and central executive component to manipulate the information. Some studies have found evidence for different patterns of functional connectivity during different working memory task conditions (e.g., as a function of load, stimulus type, or task phase).  Such ﬁndings suggest that functional connectivity changes could reﬂect differences in task engagement or responsivity of brain networks to modulation,rather than stable changes that persist across all task states. Some studies have found both impairments in the modality-specific perception component and the later manipulation component. In a recent meta analysis, the author draw attention to a consistent and restricted "core network" emerged form conjunctions across analyses of specific task designs and contrasts. This distributed network was believed to be active in WM task ignoring the task type , stimulus type and WM load and may be act as a base part in WM.

Schizophrenia has often been conceived as a disorder of connectivity between components of large-scale brain networks(Lynall, 2010). A few studies has focused on the relationship between WM impairment and disturbed functional connectivity both in the resting state and under various task. Some imagining studies reporting altered 
patterns of interregional functional connectivity in patients with schizophrenia during working memory task performance (Meyer-Lindenberg et al., 2001; Quintana et al., 2003; Schlosser et al., 2003a; Whalley et al., 2005). However, seldomly studies have investigate the functional connectivity in resting state.

@@

The resting state fMRI data presented here was collected from six hospitals in China which paticipated in the Brainnetome Project for Schizophrenia. The six hospitals are  Patients and controls are matched as much as possible for age, sex, handedness, and race distributions with each site. Peking University Sixth Hospital (PKUH6); Beijing Huilongguan Hospital (HLG); Xijing Hospital (XJ); Henan Mental Hospital (HM);  Renmin Hospital of Wuhan University (RWU); and Zhumadian Psychiatric Hospital (ZMD). Henan Mental Hospital provided two distinct MRI scanners: Siemens (HMS) and General Electric (HMG), for a total of eight scanning centers. The study at each center was approved by the local ethical review board. All the participants provided written informed consent.

@@

Two types of 3 T MRI scanners (four Siemens, three General Electric) were used at the participating centers. To ensure equivalent data acquisition protocol and high quality imaging data, the scanning parameters of the functional scans at each of the six centers were set up by an experienced researcher before data acquisition. An echo planar imaging sequence was used to obtain the
functional images, the parameters were as follows: 30 axial slices, TR = 2000 ms, TE = 30 ms, matrix = 64 × 64, flip angle = 90°, FOV = 240*240 mm 2, slice thickness = 4 mm, gap = 0.4 mm. A total of 250 brain volumes were collected, resulting in a total scan time of 500 s.The MRI scan sequences and parameters for each center are listed in eTable 1. To be noticed, the time point number of the images from ZMD is 180 and is different from the 240 of the other sites. We did not exclude this site since we are eager to keep as much data and sites as possible, to support the validation of the muli-site analysis.

The iamges were preprocessed with a based in-house software: Brainnetome Toolkit which utilized Statistical Parametric Mapping SPM8 (http://www.fil.ion.ucl.ac.uk/spm). The pipeline includes the following steps: The first ten images were deleted for the signal equilibration. The remaining iamges were conducted for slice acquisition correction and head motion correction.  The fMRI data which had
less than 3.0 mm of head motion and 3.0° of angular rotation were included.  Moreover, the mean framewise displacement (FD) was computed by averaging FDi from every time point for each subject. There were no differences for the mean FD between groups
( t = 0.413, p = 0.682) (Table 1). Then the fMRI images were normalized to the standard Montreal Neurological Institute (MNI) template provided by SPM and resample to the 3-mm isotropic voxels. Artefacts due to changes in global, ventricle and white matter signals, residual motion were removed using voxel-wise regression. A temporal filter (0.01 Hz < f < 0.08 Hz) were used to reduce the low-frequency drift and physiological high frequency respiratory and cardiac noise.  Finally, the data was smoothed with an isotropic Gaussian kernel of 6 mm full-width at half-maximum.

@@

Working memory is the result of various combinations of processes, no processes ((and correspondingly no brain structures)are unique or speci?c to working memory. ( Eriksson, 2015)Many brain regions interact during working memory andinclude "executive" regions in the PFC, parietal cortex, and basal ganglia, as well as regions specialized for processing the particular representations to be maintained, such as the fusiform face area for maintaining face information. Persistent neural activity in various brain regions accompanies working memory and is functionally necessary for maintenance and integration of information in working memory.

@@

The inferior frontal gyrus/anterior insula(IFG/AI) was suggested to be involved in elaborate attentional and working memory processing,  (Mattie Tops,2011). Some evidence has suggsted the IFG/AI might involved in cognitive control in working memory tasks. These ventral corticolimbic control pathways that include the IFG/AI , may adapt to working memory context that differ in the level of predictability. Anohter within subject study find the left and right IFG showed a conjunction between working memory and inhibition tasks within subjects, which indicate some component of excutive function may interactivate with the working memory systerm in working memory tasks (MacNab, 2008).

@@

A large amount of studies focused on working memory in schizophrenia, most of these are by using task fmri.  The complex pattern of hyper- and hypoactivation consistently found across studies implies that rather than focusing on DLPFC dysregulation, researchers should consider the entire network of regions involved in a given task when making inferences about the biological mechanisms of schizophrenia (David C. Glahn, 2005).

>>>>

Working memory is typically defined as the ability to maintain and manipulate information over short periods of time. It is widely agreed that WM involves several different component processes. Working memory is a temporal process to maintain and manipulate information over short periods of time. WM has been shown to be impaired in many neurological and psychiatric syndromes including Schizophrenia. It was thought to be tightly associated with cognitive deficits in Schizophrenia (Green 2000). WM itself is a complex process which has been multidimensionally related to the psychosis. A large body of work has demonstrated that individuals with schizophrenia have deficits in working memory (Goldman-Rakic, 1991; Park and Holzman, 1992). This study may shed light on the treatment of Schizophrenia.

Thus, an examination of the underlying neurological mechanism provided us with insight into the causes, progression and even treatment of schizophrenia. There has been abundant imaging studies in this area, but the differences among the tasks and material make it difficult to make a conclusion. There have been evidence of multi-modality deficits: visual spatial, visual object, verbal, and other types of working memory. One popular and original model of WM is a system encoding and maintaining modality-specific short-term memory and central executive component to manipulate the information. Some studies have found evidence for different patterns of functional connectivity during different working memory task conditions (e.g., as a function of load, stimulus type, or task phase).  Such findings suggest that functional connectivity changes could reflect differences in task engagement or responsivity of brain networks to modulation,rather than stable changes that persist across all task states. Some studies have found both impairments in the modality-specific perception component and the later manipulation component. In a recent meta analysis, the author draw attention to a consistent and restricted "core network" emerged from conjunctions across analyses of specific task designs and contrasts. This distributed network was believed to be active in WM task ignoring the task type , stimulus type and WM load and may be act as a base part in WM.

Schizophrenia has often been conceived as a disorder of connectivity between components of large-scale brain networks(Lynall, 2010). A growing number of studies have reported altered functional connectivity in schizophrenia during putatively “task-free” states and during the performance of cognitive tasks. (Grega Repovs 2012)  A few studies has focused on the relationship between WM impairment and disturbed functional connectivity both in the resting state and under various task. The connectivity could be modulated by the task demands. Some imagining studies reporting altered patterns of interregional functional connectivity in patients with schizophrenia during working memory task performance (Meyer-Lindenberg et al., 2001; Quintana et al., 2003; Schlosser et al., 2003a; Whalley et al., 2005). However, seldomly studies have investigate the functional connectivity in resting state. Connectivity within the DMN and FP have been found significantly different between resting state and 0-back, and was further modulated by memory load. (Grega Repovs 2012) Yet most of the existing find are during working memory tasks, we are still eager to find wheather we could detect functional connectivity impairments in resting state, which may be a potential trait marker of WM impairment in schizophrenia patients.

@@

The resting state fMRI data presented here was collected from six hospitals in China which participated in the Brainnetome Project for Schizophrenia. The six hospitals are  Patients and controls are matched as much as possible for age, sex, handedness, and race distributions with each site. Peking University Sixth Hospital (PKUH6); Beijing Huilongguan Hospital (HLG); Xijing Hospital (XJ); Henan Mental Hospital (HM);  Renmin Hospital of Wuhan University (RWU); and Zhumadian Psychiatric Hospital (ZMD). Henan Mental Hospital provided two distinct MRI scanners: Siemens (HMS) and General Electric (HMG), for a total of eight scanning centers. The study at each center was approved by the local ethical review board. All the participants provided written informed consent.

@@

Two types of 3 T MRI scanners (four Siemens, three General Electric) were used at the participating centers. To ensure equivalent data acquisition protocol and high quality imaging data, the scanning parameters of the functional scans at each of the six centers were set up by an experienced researcher before data acquisition. An echo planar imaging sequence was used to obtain the
functional images, the parameters were as follows: 30 axial slices, TR = 2000 ms, TE = 30 ms, matrix = 64 × 64, flip angle = 90°, FOV = 240*240 mm 2, slice thickness = 4 mm, gap = 0.4 mm. A total of 250 brain volumes were collected, resulting in a total scan time of 500 s.The MRI scan sequences and parameters for each center are listed in eTable 1. To be noticed, the time point number of the images from ZMD is 180 and is different from the 240 of the other sites. We did not exclude this site since we are eager to keep as much data and sites as possible, to support the validation of the multi-site analysis.

时间校正，头动校正，配准，降噪，滤波，空间平滑

The iamges were preprocessed with a based in-house software: Brainnetome Toolkit which utilized Statistical Parametric Mapping SPM8 (http://www.fil.ion.ucl.ac.uk/spm). The pipeline includes the following steps: The first ten images were deleted for the signal equilibration. The remaining iamges were conducted for slice acquisition correction and head motion correction.  The fMRI data which had
less than 3.0 mm of head motion and 3.0° of angular rotation were included.  Moreover, the mean frame-wise displacement (FD) was computed by averaging FDi from every time point for each subject. There were no differences for the mean FD between groups
( t = 0.413, p = 0.682) (Table 1). Then the fMRI images were normalized to the standard Montreal Neurological Institute (MNI) template provided by SPM and resample to the 3-mm isotropic voxels. Artefacts due to changes in global, ventricle and white matter signals, residual motion were removed using voxel-wise regression. A temporal filter (0.01 Hz < f < 0.08 Hz) were used to reduce the low-frequency drift and physiological high frequency respiratory and cardiac noise.  Finally, the data was smoothed with an isotropic Gaussian kernel of 6 mm full-width at half-maximum.

Functional connectivity analysis

Regions of interest (ROIs)

\subsection*{fcMRI Analyses}

To perform the fcMRI analyses, time series from the resting-state scan were extracted by averaging the time series of all voxels in each ROI. The resulting time series of each ROI were then entered into the following connectivity analyses.
The Pearson (full) correlation coefficient was used to estimate functional connectivity between each pair of time series. A Fisher Z transform

\subsection*{Statistical analysis}

Group differences in the basic demographics at each center were examined with two-tailed t-tests and chi-square tests using PASW Statistics 18.0.

For every single site separately, the diagnose effect on each FC value was assessed using a two-sample t-test comparing images between SZ and NC from each site. A second level random effects meta analysis was used to pooling the single site statistics, in order to assess the replication of the effects from the entire dataset. The effect sizes were calculated using Hedge’s g, which provides an unbiased standardized mean difference that incorporates a correction for small sample sizes. Hedge’s g was calculated using each center’s t-statistic and the corresponding sample size. All significance thresholds were set to p < 0.001 FDR-corrected for multiple comparison. The meta-analysis was conducted in R with meta-for packages.

We then performed association analysis for those fc connectivities that passed the significant test. We then examined the relationship between FC values of those that passed the significant test and demographic variables in patients.  Linear regression was then conducted to determine a possible association between those significantly FC connectivity and clinical scores including (PANSS positive , PANSS negative, PANSS general and PANSS total). These analyses are conducted both separately in each site and in the whole data. In whole data analysis, the site factor was included in the GLM model. An additional analyses were also performed with covariate of age and sex.

\section*{Results}

Demographics and clinical data

Sociodemographic and psychopathological data are presented by center in Table 1. No statistically significant differences in age and sex were noted between the schizophrenia patients and HCs at each center. The patients had significantly fewer years of education than the HCs in six out of eight centers.

Group comparisons of FC connectivities at each center

Mixed effect multicenter analysis

Compared to the normal group, four FC conncevities was reduced in the patients: they are the connection between the left IFG/caudal lateral prefrontal gyrus and the left anterior insula, Left IFG/caudal lateral prefrontal gyrus and the Left inferior frontal gyrus pars opercularis, Right intraparietal sulcus and the Left inferior frontal gyrus pars opercularis, Right intraparietal sulcus and the Right inferior frontal gyrus pars opercularis. No functional connectivity was detected with a enhanced effect in the patients group.

@@

Working memory is the result of various combinations of processes, no processes ((and correspondingly no brain structures)are unique or specific to working memory. ( Eriksson, 2015)Many brain regions interact during working memory and include "executive" regions in the PFC, parietal cortex, and basal ganglia, as well as regions specialized for processing the particular representations to be maintained, such as the fusiform face area for maintaining face information. Persistent neural activity in various brain regions accompanies working memory and is functionally necessary for maintenance and integration of information in working memory.

@@

The inferior frontal gyrus/anterior insula(IFG/AI) was suggested to be involved in elaborate attentional and working memory processing,  (Mattie Tops,2011). Some evidence has suggsted the IFG/AI might involved in cognitive control in working memory tasks. These ventral cortico-limbic control pathways that include the IFG/AI , may adapt to working memory context that differ in the level of predictability. Anohter within subject study find the left and right IFG showed a conjunction between working memory and inhibition tasks within subjects, which indicate some component of excutive function may interactivate with the working memory systerm in working memory tasks (MacNab, 2008). The fronto-opercular, intraparietal and anteriro cingulate cortex may form a circuit for non-articulatory maintenance of phonological information(Henseler 2009).

@@

A large amount of studies focused on working memory in schizophrenia, most of these are by using task fmri.  The complex pattern of hyper- and hypoactivation consistently found across studies implies that rather than focusing on DLPFC dysregulation, researchers should consider the entire network of regions involved in a given task when making inferences about the biological mechanisms of schizophrenia (David C. Glahn, 2005). A PET study has found impaired interaction between right lateral prefrontal cortex and bilateral inferior parietal region in SZ patients compared with normal patients during working memory processing(Jae-Jin Kim, 2003). There has been research reporting the disconneciton between ...
